Verifying Demonstrations

### 2020-03-04 paper/DB-Schnorr-UWB.tex
Starts adapting UWB-AKE for DAA/CL04
Reorders UWB again, improves DBSHW overview

We will first present our version of the \ac{DB} Schnorr protocol for the 
mutual-trust case, where the prover and verifier have a pre-shared secret.
Then we will adapt it by introducing \iac{AKE} protocol.
The authentication in this \ac{AKE} will leverage that our protocol is \iac{DB} 
\ac{ZKPK} for discrete logarithms.
If the \ac{AKE} is based on discrete logarithms we can prove the distance bound 
while we authenticate the key.
This allows us to embed this protocol into tamper resistant hardware to remove 
the mutual-trust requirement.

@@

We will now introduce an extended protocol, \(\DBSHW\), which represents the 
secure hardware implementation.
The \ac{PKI} is based on CL04~\cite{CLsignatures} signatures.
\Cref{DBSHW-overview} presents an overview.

    \multicolumn{2}{l}{%
      Static \(\hat g, \hat q, \sk_P\in \ZZ_q, \pk_P = \hat g^{\sk_P}, \sigma\)
    }
    & Static \(\hat g, \hat q\)
    \\

    \midrule
    \multicolumn{3}{c}{\textbf{Pre-setup, authenticated key agreement}} \\

    \(x\rgets \ZZ_q\)
    & \(\ProtoSendRight{g^x}\)
    & \(k \gets \pk_P^{\sk_V}\)
    \\

    \(k \gets \pk_V^{\sk_P}\)
    & \(\ProtoSendLeft{\pk_V}\)
    & \(\pk_V\)
    \\

    \(\rho, \hat\rho \rgets \ZZ_q, R\gets g^{\rho}, \hat R\gets g^{\hat\rho}\)
    & \(\ProtoSendRight{R, \hat R}\)
    & \(\hat A\gets \pk_P\)
    \\

    \(\hat s_1\gets \hat\rho - c_1\sk_P \pmod q\)
    &
    &
    \\

    \(\hat s_m\gets \hat\rho - c_m\sk_P \pmod q\)
    &
    &
    \\

  \end{tabular}
  \caption{%
    One-round protocol instance of the \(\DBSHW\) protocol instantiating 
    \(\PK[\alpha][A = g^\alpha]\).
    Each transmission (arrow in the diagram) uses \(\UWBPR\).
    The protocol actually proves \(\PK[\alpha, \sk_P][A = g^\alpha \land
      \hat A = \hat g^{\sk_P}]\).
    This is for the device to prove knowledge of the key~\(\sk_P\) to prevent 
    \ac{DBTF}.
    The protocol should be repeated \(n\) times to achieve the desired 
    soundness and distance-bounding errors.
  }%
  \label{DBSHW-overview}
\end{figure*}

The \(\DBSHW\) secure hardware device must contain a (static) private 
key~\(\sk_P\) for the prover with a digital certificate~\(\cert_P\) (and 
\(\cert_V\)) for the public key~\(\pk_P\) (and \(\pk_V\)).

\paragraph*{Pre-setup}

The pre-setup phase consists of an authenticated key agreement.
We use the static \ac{DHKE} to simplify the explanation, any authenticated key 
agreement protocol based on discrete logarithms can be used.
The prover and verifier exchange certificates, \(\cert_P\) and \(\cert_V\), and 
compute the agreed key~\(k = \pk_V^{\sk_P} = \pk_P^{\sk_V}\).
If the prover's certificate is invalid, the protocol aborts.

The setup phase is similar to \(\DBS\) (\cref{SchnorrFigure}).
\(\DBS\) performs \(\PK[\alpha][A = g^\alpha]\) whereas \(\DBSHW\) must perform 
\(\PK[\alpha, \sk_P][\hat A = g^\alpha \land \hat g^{\sk_P}]\).
They use the agreed upon key~\(k\) as the key for \(\UWBPR\).

@@

\subsection{\acs*{DBDF} resistance}

This protocol uses the \ac{DBMF} resistance of \(\DBS\) to ensure \ac{DBDF} 
resistance against a malicious prover.
The trusted hardware uses the authenticated key agreement to establish the 
shared secret \(k\), thus a verifier prevents a malicious prover from learning 
\(k\) --- the main limit of~\cite{UWBPR}.

The hardware device (\DBSHW) could commit \ac{DBTF}\footnote{%
  Arguably, this should not be an issue since the device is trusted.
  However, we include this construction because it is a neat way to provide 
  \ac{DBTF} resistance for this part of the protocol as well.
  To not have this feature, one simply adds the authenticated key exchange step 
  to \cref{SchnorrFigure}.
} by giving the key~\(k\) to a malicious prover --- this would allow that 
prover to authenticate once.
To prevent this possibility we include a proof of knowledge of the device's 
private key~\(\sk_P\), \ie that \(\pk_P = \hat g^{\sk_P}\).
This way, an accomplice can only pass if he is given both \(k\) and \(\sk_P\).

>>>>

We will first present our version of the \ac{DB} Schnorr protocol for the 
mutual-trust case, where the prover and verifier have a pre-shared secret.
Then we will adapt it by introducing \iac{PKI}.
We will use this \ac{PKI} to create \iac{AKE} which leverages that our protocol 
is \iac{DB} \ac{ZKPK} for discrete logarithms.
If the \ac{PKI} is based on discrete logarithms we can prove the distance bound 
while we authenticate the key.
This turns the mutual-trust by shared-key into mutual trust by 
\ac{PKI}.

@@

@@

We will now introduce an extended protocol, \(\DBSHW\), which represents the 
secure hardware implementation.
\Cref{DBSHW-overview} presents an overview of our protocol.
For simplicity of exposition, we present a \ac{PKI} based on plain 
CL04~\cite{CLsignatures} signatures.
However, CL04 is the base for the \ac{DAA} protocol, which is used for 
\ac{TPM}, so our approach can make \ac{DAA} \acl{DB}.

In summary, the difference is as follows.
\(\DBS\) has a shared key~\(k\) and performs \(
  \PK[\alpha][A = g^\alpha]
\) with a distance bound.
\(\DBSHW\) first runs \iac{DHKE} to establish the shared key~\(k =
\hat g^{xy}\) from \(X = \hat g^x, Y = \hat g^y\), then performs \[
  \PK[\alpha, x, \sk_P, \sigma][
    A = g^\alpha \land
    X = \hat g^x \land
    \Blind[\sigma] = \Sign[_\sk][\sk_P]
  ],
\] also with a distance bound; where \(\sk_P\) is the prover's private key, 
\(\sigma\) is a signature issued with signing key~\(sk\) and \(\Blind\) 
reblinds a signature.
Proving knowledge of \(x\) prevents any \ac{MITM} attack against the \ac{DHKE}.
Proving knowledge of a signature~\(\sigma\) by \iac{CA} on the private 
key~\(\sk_P\) ensures that the prover is a device in the \ac{PKI} of the 
\ac{CA}.

\Cref{DBSHW-overview} provides an overview.

    \multicolumn{2}{l}{%
      Static \(\sk_P\in \ZZ_q, \sigma = (a, A, b, B, c)\)
    }
    & Static \(\pk\)
    \\

    \midrule
    \multicolumn{3}{c}{\textbf{Pre-setup key-agreement}} \\

    \(x\rgets \ZZ_q\)
    & \(\ProtoSendRight{g^x}\)
    & \(y\rgets \ZZ_q\)
    \\

    \(k \gets (g^y)^x\)
    & \(\ProtoSendLeft{g^y}\)
    & \(k \gets (g^{x})^y\)
    \\

    \(\rho, \hat\rho \rgets \ZZ_q, R\gets g^{\rho}, \hat R\gets g^{\hat\rho}\)
    & \(\ProtoSendRight{R, \hat R}\)
    & \(\hat A\gets g^x\)
    \\

    \(\hat s_1\gets \hat\rho - c_1 x \pmod q\)
    &
    &
    \\

    \(\hat s_m\gets \hat\rho - c_m x \pmod q\)
    &
    &
    \\

  \end{tabular}
  \caption{%
    One-round protocol instance of the \(\DBSHW\) protocol instantiating 
    \(\PK[\alpha][A = g^\alpha]\).
    Each transmission (arrow in the diagram) uses \(\UWBPR\) (except in the 
    pre-setup phase).
    The protocol actually proves \(\PK[\alpha, x][A = g^\alpha \land
      \hat A = \hat g^x]\).
    This is ties \(x\), and consequently \(k\), to \(\alpha\).
    The protocol (setup, distance-bounding and verification phases) should be 
    repeated \(n\) times to achieve the desired soundness and distance-bounding 
    errors.
  }%
  \label{DBSHW-overview}
\end{figure*}

\paragraph*{One-time initialization}

The prover's \(\DBSHW\) device must be initialized with a (static) private 
key~\(\sk_P\) and a signature~\(\sigma\) on that key by some \ac{CA} with 
public key~\(\pk\) and signing key~\(\sk\).
These are computed as follows~\cite[cf.][Sect.~4.2]{CLsignatures}.

We have public parameters \(q, G, g, \G, \g, e\); where \(e\colon G\times G\to 
\G\) is a bilinear map, \(g\in G\) and \(\g\in \G\) are generators of prime 
order \(q\) (hence \(e(g, g) = \g\)).

We let the \acg{CA} signing key be \(\sk = (x, y, z)\rgets \ZZ_q^3\) and the 
public key be \(\pk = (q, G, \G, g, \g, e, X, Y, Z)\), where \(X = g^x, Y = 
g^y, Z = g^z\).

The prover chooses \(\sk_P\rgets \ZZ_q\) and computes a commitment \(M = 
g^{\sk_P} Z^r\), where \(r\rgets \ZZ_q\).
Now the prover convinces the \ac{CA} that he knows the key by running the 
following \ac{ZKPK} protocol: \(
  \PK[\sk_P, r][M = g^{\sk_P} Z^r].
\)

Now the \ac{CA} computes the signature~\(\sigma = (a, A, b, B, c)\), where
\begin{itemize}
  \item \(a\gets g^\alpha\), where \(\alpha\rgets \ZZ_q\),
  \item \(A\gets g^z, b\gets a^y, B\gets A^y\),
  \item \(c\gets a^x M^{\alpha x y}\).
\end{itemize}
and gives to the prover.

\paragraph*{Pre-setup key-agreement}

The pre-setup phase consists of a key agreement, the prover and verifier must 
agree on a shared key~\(k\) to use for the \(\UWBPR\) protocol on the physical 
layer.

To establish \(k\), we use the \ac{DHKE}.
The prover chooses \(x\rgets \ZZ_q\), the verifier chooses \(y\rgets \ZZ_q\) 
and they exchange \(X = \hat g^x, Y = \hat g^y\) and both compute
\(k = \hat g^{xy}\), for some generator \(\hat g\) in \iac{DH} group of prime 
order \(q\).

We will see later that if the adversary attempted \iac{MITM} attack, our 
\(\DBSHW\) protocol will detect this, because the adversary will fail 
\iac{ZKPK}.

The setup phase is similar to \(\DBS\) (\cref{SchnorrFigure}).
They use the agreed upon key~\(k = g^{xy}\) as the key for \(\UWBPR\).
\(\DBS\) performs \(
  \PK[\alpha][A = g^\alpha]
\) (with a distance bound) whereas \(\DBSHW\) must perform \(
  \PK[\alpha, x][\hat A = g^\alpha \land X = \hat g^x]
\) (also with a distance bound).
